Emulating Comparative Oncology Trials with Real-world Evidence Studies (ENCORE): Process Development and Methodological Considerations for Oncology Real-World Data

### 2025-07-06 manuscript/manuscript.tex
Manuscript updated

\ul{\textbf{Manuscript word count:}} 4,509 words / 8,000 words

\ul{\textbf{Abstract word count:}} 224 words / 250 words

@@

\ul{\textbf{\emph{Proposed target journals:}}} \emph{CPT =\textgreater{}
JCO CCI =\textgreater{} \ldots{}}

\emph{Manuscript last updated: 2025-07-06 14:54:57.141696}

@@

Randomized controlled trials (RCTs) are the evidentiary gold standard
methodology for establishing the efficacy and safety of medical
products. Guided by the 21\textsuperscript{st} Century Cures
Act,\textsuperscript{1} the Food and Drug Administration (FDA)
established a framework to factor considerations for use of real-world
evidence (RWE) generated from real-world data (RWD) such as electronic
health records (EHR) to support approval of new indications or satisfy
postmarketing requirements.\textsuperscript{1} Accounting for
approximately 30\% of new FDA drug approvals, oncology was the disease
area with the most new approvals in 2024\textsuperscript{2} as well as
several indication expansions, and has many areas of high unmet medical
need to treat serious conditions. Therefore, RWE can have a particularly
important potential to complement evidence from RCTs in the field of
oncology. Potential uses include the assessment of effectiveness in
specific patient populations that are not adequately represented in
RCTs, or the precision oncology-focused discovery of biomarkers among
pan-tumor populations that harbor specific genomic and
immuno-pathological signatures.

However, the comparability and transportability of results derived
between observational studies and RCTs depends on multiple factors.
Frequently referenced limitations include lack of baseline randomization
and the imbalance in prognostic factors resulting thereof, missing data,
unmeasured confounding, small study sizes, data
discontinuity,\textsuperscript{3, 4} adoption of changes in treatment
guidelines over time in routine clinical care and the inability to
measure and emulate common eligibility criteria, including prognostic
factors, and standardized response assessments in real-world data
(RWD).\textsuperscript{5} While examples of oncology trial emulations
have been published,\textsuperscript{5--7} a systematic and scaled
approach to emulate a diverse set of oncology trials with multiple
heterogeneous databases is necessary to increase confidence in the
interpretability of observational studies, evaluate regulatory
considerations, and to provide context as to which questions can be
validly answered.

@@

The \emph{Emulation of Comparative Oncology Trials with Real-world
Evidence} (ENCORE) project\textsuperscript{10} aims to extend this work
to oncology. Clinical studies in oncology come with unique
methodological challenges which must be systematically explored and
understood. Building on a process co-developed with the FDA through RCT
DUPLICATE,\textsuperscript{8} this expansion to oncology will emulate 12
randomized clinical trials using multiple specialty oncology EHR data
sources. The process will emphasize transparency and include a data
fitness-for-use assessment\textsuperscript{11, 12} for each RWD source
with respect to each trial emulation as well as extensive sensitivity
analyses to assess robustness of findings.

@@

Although there have been substantial methodological advancements to
increase our understanding on the emulation and comparison of real-world
progression-free survival (PFS) and objective response rates (ORR) to a
RECISTv1.1\textsuperscript{14}-based PFS and ORR assessment in
RCTs\textsuperscript{15, 16}, imaging-based evaluations still require a
level of granularity which is not well reflected in chart-abstracted
assessments of a patient's progression in routine
care.\textsuperscript{17, 18} The timing and cadence of intervals
between progression assessments can differ between RCTs and routine care
which may result in measurement error and bias.\textsuperscript{19}
Given this, and the large number of other methodological challenges
including missing data, small sample sizes, data discontinuity and
rapidly changing guideline treatments, we focus on the emulation of
overall survival (OS) as the endpoint of interest. Therefore, we include
only trials that have reported OS as a pre-specified endpoint in the
protocol.

@@

For each selected RCT, a detailed protocol, pre-specifying key elements
of the trial emulation, will be developed using the HARPER protocol
template\textsuperscript{21, 22} and will be registered on
ClinicalTrials.gov after review by an expert panel comprising clinicians
and FDA reveiwers. Following the target trial emulation framework, we
will provide an explicit description and rationale on how each element
will be emulated including database selection, covariate measurement,
operationalization of key eligibility criteria, study design, data
analysis and causal contrasts of interest.\textsuperscript{23, 24} Since
it is common that oncology RCTs update survival estimates periodically
based on accrued follow-up time, the protocol will give a brief summary
of each emulated RCT and specify which target OS estimates will be used
for agreement metric evaluation (see
Section~\ref{sec-agreement-metrics}). All eligibility criteria will be
extracted based on publications, publicly available protocols, and
statistical analysis plans of the selected RCT.

@@

To formally compare treatment effects between RCTs and their respective
emulations, we will adapt the approach of the RCT-DUPLICATE
project.\textsuperscript{8, 30} That is, for the primary endpoint of
interest (hazard ratio {[}HR{]} for OS and corresponding 95\% confidence
intervals), we will derive three qualitative agreement metrics:
statistical agreement, estimate agreement and agreement based on the
standardized mean difference (SMD). Examples are illustrated in
Table~\ref{tbl-metrics}.

@@

To establish an analytic cohort, missingness will be assessed across
patients who meet all eligibility criteria, including those with missing
values as a first iteration. These missing data investigations will
empirically assess assumptions on potential underlying missingness
mechanisms according to Rubin's classification of missing data (i.e.,
missing completely at random {[}MCAR{]}, missing at random {[}MAR{]} and
missing not at random {[}MNAR{]}).\textsuperscript{32} We will adopt a
principled process on missing data that empirically evaluates different
aspects across partially observed covariates based on three group
diagnostics.\textsuperscript{33, 34} These diagnostics cover (1)
comparisons of patients characteristics with and without an observed
level of the partially observed covariate, (2) ability to predict
missingness given observed data, and (3) assessments if outcomes between
patients with a missing value are systematically different. Expert
domain knowledge and assumptions about the underlying missing data
structure through canonical causal diagrams\textsuperscript{35} will
additionally inform decisions regarding the inclusion or exclusion of
individuals with missing values in key eligibility criteria and
potential sensitivity analyses to assess the robustness of these
decisions.\textsuperscript{36}

@@

For the estimation of marginal treatment effects, we will employ
propensity score methods to adjust for measured confounding between
treatment arms. The selection of relevant prognostic covariates will be
based on expert clinical knowledge and published literature on
prognostic scores in oncology.\textsuperscript{41} The implementation of
propensity scores in combination with multiple imputation will follow
the `\emph{within}' methodology as described by Leyrat et
al.\textsuperscript{42, 43} As appropriate, propensity score matching or
weighting will be applied to each imputed dataset. The marginal
treatment effect will then be estimated in each imputed and matched or
weighted dataset separately and pooled into a final estimate following
Rubin's rule.\textsuperscript{44, 45} This approach has been shown to
lead to unbiased estimates across different simulated scenarios with a
sufficient estimation of the variance.\textsuperscript{42}

@@

Similarly, survival probabilities for individual time points will be
estimated in each imputed and propensity score matched or weighted
dataset according to the Kaplan-Meier method.\textsuperscript{40} Since
survival probabilities typically do not follow normal distributions
which are required to apply Rubin's rule, these will be transformed
through a complementary log-log transformation \(log(-log(1-pr(surv)))\)
with \(pr(surv)\) denoting the survival probability at a given follow-up
time during.\textsuperscript{49, 50} The transformed survival
probabilities are then pooled across imputed datasets and individual
time points following Rubin's rule and back-transformed via
\(1-exp(-exp(qbar))\) with \(qbar\) denoting the pooled survival
probability. The median survival time can be finally determined by
extracting the time point during follow-up at which the survival
probability drops below 0.5 for the first time.

@@

Building on an established approach developed by the RCT DUPLICATE
project\textsuperscript{8, 30}, the ENCORE project aims to emulate 12
oncology trials using four oncology EHR data sources to inform the
potential use of RWD for regulatory purposes in oncology. The project
focuses on four common cancers and will evaluate the agreement of
treatment effect estimates between RCTs and their respective emulations.
Historically, administrative claims databases have been the predominant
data source used to evaluate medical product safety and effectiveness in
the postmarketing setting. With increasing access to EHR data and
availability of granular clinical data elements often necessary to
gather fit-for-use data in oncology, as well as a maturing set of
methodological approaches for causal inference using such data, the
ENCORE project is one of the first of its kind to evaluate contexts
where RWD may be reliably used to draw similar conclusions compared to
RCTs in the field of oncology.

The RCT DUPLICATE initiative\textsuperscript{8, 30} identified multiple
emulation challenges that may be similar in ENCORE. One particular
aspect that we may not always be able to emulate is the exact
distribution of patient characteristics of the trial population. This
can be due to the lack of data granularity and comprehensiveness to
emulate relevant eligibility criteria or the fact that large pivotal
trials in oncology are typically conducted in multiple countries
worldwide (all considered databases reflect the US only). This may
impact the results of this trial emulation given that the distribution
of genetic alterations and biomarkers, prognosis, and factors that drive
heterogeneous treatment outcomes of certain cancers can differ between
countries. For example, certain cancers (e.g., GI cancers) or certain
genetic mutations (e.g., EGFR mutations in lung cancer) are much more
prevalent in certain geographic regions and populations compared to the
US. Additionally, treatment landscapes such as the preference for
specific regimens or compounds, their use in earlier or later lines of
therapies, and supportive care practices can differ between geographic
regions, too.

@@

\phantomsection\label{refs}
\begin{CSLReferences}{0}{1}
\bibitem[\citeproctext]{ref-RWEFDA}
\textbf{1}. Framework for FDA's real-world evidence program (last
accessed 11/28/2024) {[}Internet{]}, 2018Available from:
\url{https://www.fda.gov/media/120060/download?attachment}

\bibitem[\citeproctext]{ref-mullard20242023}
\textbf{2}. Mullard A: 2023 FDA approvals. Nat Rev Drug Discov
23:88--95, 2024

\bibitem[\citeproctext]{ref-Merola2022}
\textbf{3}. Merola D, Schneeweiss S, Schrag D, et al: An algorithm to
predict data completeness in oncology electronic medical records for
comparative effectiveness research {[}Internet{]}. Annals of
Epidemiology 76:143--149, 2022Available from:
\url{http://dx.doi.org/10.1016/j.annepidem.2022.07.007}

\bibitem[\citeproctext]{ref-joshua2022longitudinal}
\textbf{4}. Joshua Lin K, Jin Y, Gagne J, et al: Longitudinal data
discontinuity in electronic health records and consequences for
medication effectiveness studies. Clinical Pharmacology \& Therapeutics
111:243--251, 2022

\bibitem[\citeproctext]{ref-rider2024emulations}
\textbf{5}. Rider JR, Wasserman A, Slipski L, et al: Emulations of
oncology trials using real-world data: A systematic literature review.
American journal of epidemiology kwae346, 2024

\bibitem[\citeproctext]{ref-merola2023aetion}
\textbf{6}. Merola D, Campbell U, Gautam N, et al: The aetion coalition
to advance real-world evidence through randomized controlled trial
emulation initiative: oncology. Clinical Pharmacology \& Therapeutics
113:1217--1222, 2023

\bibitem[\citeproctext]{ref-merola2024calibrating}
\textbf{7}. Merola D, Campbell U, Lenis D, et al: Calibrating
observational health record data against a randomized trial. JAMA
Network Open 7:e2436535--e2436535, 2024

\bibitem[\citeproctext]{ref-wang2023emulation}
\textbf{8}. Wang SV, Schneeweiss S, Franklin JM, et al:
\href{https://doi.org/10.1001/jama.2023.4221}{Emulation of randomized
clinical trials with nonrandomized database analyses: Results of 32
clinical trials}. Jama 329:1376--1385, 2023

\bibitem[\citeproctext]{ref-heyard2024design}
\textbf{9}. Heyard R, Held L, Schneeweiss S, et al: Design differences
and variation in results between randomised trials and non-randomised
emulations: Meta-analysis of RCT-DUPLICATE data {[}Internet{]}. BMJ
medicine 3, 2024Available from:
\url{https://doi.org/10.1136/bmjmed-2023-000709}

\bibitem[\citeproctext]{ref-encoreFDA}
\textbf{10}. Calibrating real-world evidence studies in oncology against
randomized trials: ENCORE (last accessed 11/28/2024) {[}Internet{]},
2024Available from:
\url{https://www.fda.gov/about-fda/oncology-center-excellence/calibrating-real-world-evidence-studies-oncology-against-randomized-trials-encore}

\bibitem[\citeproctext]{ref-rivera2024oncology}
\textbf{11}. Rivera DR, Eckert JC, Rodriguez-Watson C, et al: The
oncology QCARD initiative: Fostering efficient evaluation of initial
real-world data proposals. Pharmacoepidemiology and Drug Safety
33:e5818, 2024

\bibitem[\citeproctext]{ref-gatto2022structured}
\textbf{12}. Gatto NM, Campbell UB, Rubinstein E, et al: The structured
process to identify fit-for-purpose data: A data feasibility assessment
framework. Clinical Pharmacology \& Therapeutics 111:122--134, 2022

\bibitem[\citeproctext]{ref-tasneem2012database}
\textbf{13}. Tasneem A, Aberle L, Ananth H, et al: The database for
aggregate analysis of ClinicalTrials. Gov (AACT) and subsequent
regrouping by clinical specialty. PloS one 7:e33677, 2012

\bibitem[\citeproctext]{ref-eisenhauer2009new}
\textbf{14}. Eisenhauer EA, Therasse P, Bogaerts J, et al: New response
evaluation criteria in solid tumours: Revised RECIST guideline (version
1.1). European journal of cancer 45:228--247, 2009

\bibitem[\citeproctext]{ref-ThanCCR}
\textbf{15}. Ton TGN, Pal N, Trinh H, et al: Replication of overall
survival, progression-free survival, and overall response in
chemotherapy arms of nonâsmall cell lung cancer trials using real-world
data {[}Internet{]}. Clinical Cancer Research 28:2844--2853,
2022Available from: \url{https://doi.org/10.1158/1078-0432.CCR-22-0471}

\bibitem[\citeproctext]{ref-mckelvey2024evaluation}
\textbf{16}. McKelvey BA, Garrett-Mayer E, Rivera DR, et al: Evaluation
of real-world tumor response derived from electronic health record data
sources: A feasibility analysis in patients with metastatic non--small
cell lung cancer treated with chemotherapy. JCO Clinical Cancer
Informatics 8:e2400091, 2024

\bibitem[\citeproctext]{ref-rwdRECIST}
\textbf{17}. Chen L, Davis R, Lee J, et al: Comparison of response from
RECIST1.1 and abstraction in real-world patients with lung cancer.
{[}Internet{]}. Journal of Clinical Oncology 41:e21194--e21194,
2023Available from:
\url{https://ascopubs.org/doi/abs/10.1200/JCO.2023.41.16_suppl.e21194}

\bibitem[\citeproctext]{ref-rivera2022friends}
\textbf{18}. Rivera DR, Henk HJ, Garrett-Mayer E, et al: The friends of
cancer research real-world data collaboration pilot 2.0: Methodological
recommendations from oncology case studies. Clinical Pharmacology \&
Therapeutics 111:283--292, 2022

\bibitem[\citeproctext]{ref-ackerman2024measurement}
\textbf{19}. Ackerman B, Gan RW, Meyer CS, et al: Measurement error and
bias in real-world oncology endpoints when constructing external control
arms. Frontiers in Drug Safety and Regulation 4:1423493, 2024

\bibitem[\citeproctext]{ref-nikolakopoulou2014interpret}
\textbf{20}. Nikolakopoulou A, Mavridis D, Salanti G: How to interpret
meta-analysis models: Fixed effect and random effects meta-analyses. BMJ
Ment Health 17:64--64, 2014

\bibitem[\citeproctext]{ref-wang2022harmonized}
\textbf{21}. Wang SV, Pottegård A, Crown W, et al: HARmonized protocol
template to enhance reproducibility of hypothesis evaluating real-world
evidence studies on treatment effects: A good practices report of a
joint ISPE/ISPOR task force {[}Internet{]}. Value in Health
25:1663--1672, 2022Available from:
\url{https://doi.org/10.1016/j.jval.2022.09.001}

\bibitem[\citeproctext]{ref-guidelinegeneral}
\textbf{22}. ICH guideline. General principles on plan, design and
analysis of pharmacoepidemiological studies that utilize real-world data
for safety assessment of medicines M14. Available from:
Https://database.ich.org/sites/default/files/ICH\_M14\_Step3\_DraftGuideline\_2024\_0521.pdf
(last accessed 01/07/2025)

\bibitem[\citeproctext]{ref-hernan2022target}
\textbf{23}. Hernán MA, Wang W, Leaf DE: Target trial emulation: A
framework for causal inference from observational data. Jama
328:2446--2447, 2022

\bibitem[\citeproctext]{ref-hernan2016specifying}
\textbf{24}. Hernán MA, Sauer BC, Hernández-Dı́az S, et al:
\href{https://doi.org/10.1016/j.jclinepi.2016.04.014}{Specifying a
target trial prevents immortal time bias and other self-inflicted
injuries in observational analyses}. Journal of clinical epidemiology
79:70--75, 2016

\bibitem[\citeproctext]{ref-rufibach2018}
\textbf{25}. Rufibach K: Treatment effect quantification for
time{-}to{-}event endpoints{\textendash}Estimands, analysis strategies,
and beyond {[}Internet{]}. Pharmaceutical Statistics 18:145--165,
2018Available from: \url{http://dx.doi.org/10.1002/pst.1917}

\bibitem[\citeproctext]{ref-manitz2022estimands}
\textbf{26}. Manitz J, Kan-Dobrosky N, Buchner H, et al: Estimands for
overall survival in clinical trials with treatment switching in
oncology. Pharmaceutical Statistics 21:150--162, 2022

\bibitem[\citeproctext]{ref-meyer2020open}
\textbf{27}. Meyer A-M, Davies J, Taylor M, et al: Open cohorts and
ghost-time bias in real world data, in PHARMACOEPIDEMIOLOGY AND DRUG
SAFETY. WILEY 111 RIVER ST, HOBOKEN 07030-5774, NJ USA, 2020, pp
426--426

\bibitem[\citeproctext]{ref-hernan2022causal}
\textbf{28}. Hernán MA: Causal analyses of existing databases: No power
calculations required. Journal of clinical epidemiology 144:203--205,
2022

>>>>

\ul{\textbf{Manuscript word count:}} 4,541 words / 8,000 words

\ul{\textbf{Abstract word count:}} 228 words / 250 words

@@

\emph{Manuscript last updated: 2025-07-06 16:21:00.006613}

@@

Randomized controlled trials (RCTs) are the evidentiary gold standard
methodology for establishing the efficacy and safety of medical
products. Guided by the 21\textsuperscript{st} Century Cures
Act,\textsuperscript{1} the Food and Drug Administration (FDA)
established a framework to factor considerations for use of real-world
evidence (RWE) generated from real-world data (RWD) such as electronic
health records (EHR) to support approval of new indications or satisfy
postmarketing requirements.\textsuperscript{1} Accounting for
approximately 30\% of new FDA drug approvals, oncology was the disease
area with the most new approvals in 2023\textsuperscript{2} as well as
several indication expansions, and has many areas of high unmet medical
need to treat serious conditions. Therefore, RWE can have a particularly
important potential to complement evidence from RCTs in the field of
oncology. Potential uses include the assessment of effectiveness in
specific patient populations that are not adequately represented in
RCTs, or the precision oncology-focused discovery of biomarkers among
pan-tumor populations that harbor specific genomic and
immuno-pathological signatures.

However, the comparability and transportability of results derived
between observational studies and RCTs depends on multiple factors.
Frequently referenced limitations include lack of baseline randomization
and the imbalance in prognostic factors resulting thereof, missing data,
unmeasured confounding, small study sizes, data
discontinuity,\textsuperscript{3,4} adoption of changes in treatment
guidelines over time in routine clinical care and the inability to
measure and emulate common eligibility criteria, including prognostic
factors, and standardized response assessments in real-world data
(RWD).\textsuperscript{5} While examples of oncology trial emulations
have been published,\textsuperscript{5--7} a systematic and scaled
approach to emulate a diverse set of oncology trials with multiple
heterogeneous databases is necessary to increase confidence in the
interpretability of observational studies, evaluate regulatory
considerations, and to provide context as to which questions can be
validly answered.

@@

The \emph{Emulation of Comparative Oncology Trials with Real-world
Evidence} (ENCORE) project\textsuperscript{10} aims to extend this work
to oncology. Clinical studies in oncology come with unique
methodological challenges which must be systematically explored and
understood. Building on a process co-developed with the FDA through RCT
DUPLICATE,\textsuperscript{8} this expansion to oncology will emulate 12
randomized clinical trials using multiple specialty oncology EHR data
sources. The process will emphasize transparency and include a data
fitness-for-use assessment\textsuperscript{11,12} for each RWD source
with respect to each trial emulation as well as extensive sensitivity
analyses to assess robustness of findings.

@@

Although there have been substantial methodological advancements to
increase our understanding on the emulation and comparison of real-world
progression-free survival (PFS) and objective response rates (ORR) to a
RECISTv1.1\textsuperscript{14}-based PFS and ORR assessment in
RCTs\textsuperscript{15,16}, imaging-based evaluations still require a
level of granularity which is not well reflected in chart-abstracted
assessments of a patient's progression in routine
care.\textsuperscript{17,18} The timing and cadence of intervals between
progression assessments can differ between RCTs and routine care which
may result in measurement error and bias.\textsuperscript{19} Given
this, and the large number of other methodological challenges including
missing data, small sample sizes, data discontinuity and rapidly
changing guideline treatments, we focus on the emulation of overall
survival (OS) as the endpoint of interest. Therefore, we include only
trials that have reported OS as a pre-specified endpoint in the
protocol.

@@

For each selected RCT, a detailed protocol, pre-specifying key elements
of the trial emulation, will be developed using the HARPER protocol
template\textsuperscript{21,22} and will be registered on
ClinicalTrials.gov after review by an expert panel comprising clinicians
and FDA reveiwers. Following the target trial emulation framework, we
will provide an explicit description and rationale on how each element
will be emulated including database selection, covariate measurement,
operationalization of key eligibility criteria, study design, data
analysis and causal contrasts of interest.\textsuperscript{23,24} Since
it is common that oncology RCTs update survival estimates periodically
based on accrued follow-up time, the protocol will give a brief summary
of each emulated RCT and specify which target OS estimates will be used
for agreement metric evaluation (see
Section~\ref{sec-agreement-metrics}). All eligibility criteria will be
extracted based on publications, publicly available protocols, and
statistical analysis plans of the selected RCT.

@@

To formally compare treatment effects between RCTs and their respective
emulations, we will adapt the approach of the RCT-DUPLICATE
project.\textsuperscript{8,30} That is, for the primary endpoint of
interest (hazard ratio {[}HR{]} for OS and corresponding 95\% confidence
intervals), we will derive three qualitative agreement metrics:
statistical agreement, estimate agreement and agreement based on the
standardized mean difference (SMD). Examples are illustrated in
Table~\ref{tbl-metrics}.

@@

To establish an analytic cohort, missingness will be assessed across
patients who meet all eligibility criteria, including those with missing
values as a first iteration. These missing data investigations will
empirically assess assumptions on potential underlying missingness
mechanisms according to Rubin's classification of missing data (i.e.,
missing completely at random {[}MCAR{]}, missing at random {[}MAR{]} and
missing not at random {[}MNAR{]}).\textsuperscript{32} We will adopt a
principled process on missing data that empirically evaluates different
aspects across partially observed covariates based on three group
diagnostics.\textsuperscript{33,34} These diagnostics cover (1)
comparisons of patients characteristics with and without an observed
level of the partially observed covariate, (2) ability to predict
missingness given observed data, and (3) assessments if outcomes between
patients with a missing value are systematically different. Expert
domain knowledge and assumptions about the underlying missing data
structure through canonical causal diagrams\textsuperscript{35} will
additionally inform decisions regarding the inclusion or exclusion of
individuals with missing values in key eligibility criteria and
potential sensitivity analyses to assess the robustness of these
decisions.\textsuperscript{36}

@@

For the estimation of marginal treatment effects, we will employ
propensity score methods to adjust for measured confounding between
treatment arms. The selection of relevant prognostic covariates will be
based on expert clinical knowledge and published literature on
prognostic scores in oncology.\textsuperscript{41} The implementation of
propensity scores in combination with multiple imputation will follow
the `\emph{within}' methodology as described by Leyrat et
al.\textsuperscript{42,43} As appropriate, propensity score matching or
weighting will be applied to each imputed dataset. The marginal
treatment effect will then be estimated in each imputed and matched or
weighted dataset separately and pooled into a final estimate following
Rubin's rule.\textsuperscript{44,45} This approach has been shown to
lead to unbiased estimates across different simulated scenarios with a
sufficient estimation of the variance.\textsuperscript{42}

@@

Similarly, survival probabilities for individual time points will be
estimated in each imputed and propensity score matched or weighted
dataset according to the Kaplan-Meier method.\textsuperscript{40} Since
survival probabilities typically do not follow normal distributions
which are required to apply Rubin's rule, these will be transformed
through a complementary log-log transformation \(log(-log(1-pr(surv)))\)
with \(pr(surv)\) denoting the survival probability at a given follow-up
time during.\textsuperscript{49,50} The transformed survival
probabilities are then pooled across imputed datasets and individual
time points following Rubin's rule and back-transformed via
\(1-exp(-exp(qbar))\) with \(qbar\) denoting the pooled survival
probability. The median survival time can be finally determined by
extracting the time point during follow-up at which the survival
probability drops below 0.5 for the first time.

@@

Building on an established approach developed by the RCT DUPLICATE
project\textsuperscript{8,30}, the ENCORE project aims to emulate 12
oncology trials using four oncology EHR data sources to inform the
potential use of RWD for regulatory purposes in oncology. The project
focuses on four common cancers and will evaluate the agreement of
treatment effect estimates between RCTs and their respective emulations.
Historically, administrative claims databases have been the predominant
data source used to evaluate medical product safety and effectiveness in
the postmarketing setting. With increasing access to EHR data and
availability of granular clinical data elements often necessary to
gather fit-for-use data in oncology, as well as a maturing set of
methodological approaches for causal inference using such data, the
ENCORE project is one of the first of its kind to evaluate contexts
where RWD may be reliably used to draw similar conclusions compared to
RCTs in the field of oncology.

The RCT DUPLICATE initiative\textsuperscript{8,30} identified multiple
emulation challenges that may be similar in ENCORE. One particular
aspect that we may not always be able to emulate is the exact
distribution of patient characteristics of the trial population. This
can be due to the lack of data granularity and comprehensiveness to
emulate relevant eligibility criteria or the fact that large pivotal
trials in oncology are typically conducted in multiple countries
worldwide (all considered databases reflect the US only). This may
impact the results of this trial emulation given that the distribution
of genetic alterations and biomarkers, prognosis, and factors that drive
heterogeneous treatment outcomes of certain cancers can differ between
countries. For example, certain cancers (e.g., GI cancers) or certain
genetic mutations (e.g., EGFR mutations in lung cancer) are much more
prevalent in certain geographic regions and populations compared to the
US. Additionally, treatment landscapes such as the preference for
specific regimens or compounds, their use in earlier or later lines of
therapies, and supportive care practices can differ between geographic
regions, too.

@@

\phantomsection\label{refs}
\begin{CSLReferences}{0}{0}
\bibitem[\citeproctext]{ref-RWEFDA}
\CSLLeftMargin{1. }%
\CSLRightInline{Framework for FDA's real-world evidence program (last
accessed 11/28/2024). (2018).at
\textless{}\url{https://www.fda.gov/media/120060/download?attachment}\textgreater{}}

\bibitem[\citeproctext]{ref-mullard20242023}
\CSLLeftMargin{2. }%
\CSLRightInline{Mullard, A. 2023 FDA approvals. \emph{Nat Rev Drug
Discov} \textbf{23}, 88--95 (2024).}

\bibitem[\citeproctext]{ref-Merola2022}
\CSLLeftMargin{3. }%
\CSLRightInline{Merola, D., Schneeweiss, S., Schrag, D., Lii, J. \& Lin,
K. J. \href{https://doi.org/10.1016/j.annepidem.2022.07.007}{An
algorithm to predict data completeness in oncology electronic medical
records for comparative effectiveness research}. \emph{Annals of
Epidemiology} \textbf{76}, 143--149 (2022).}

\bibitem[\citeproctext]{ref-joshua2022longitudinal}
\CSLLeftMargin{4. }%
\CSLRightInline{Joshua Lin, K. \emph{et al.} Longitudinal data
discontinuity in electronic health records and consequences for
medication effectiveness studies. \emph{Clinical Pharmacology \&
Therapeutics} \textbf{111}, 243--251 (2022).}

\bibitem[\citeproctext]{ref-rider2024emulations}
\CSLLeftMargin{5. }%
\CSLRightInline{Rider, J. R. \emph{et al.} Emulations of oncology trials
using real-world data: A systematic literature review. \emph{American
journal of epidemiology} kwae346 (2024).}

\bibitem[\citeproctext]{ref-merola2023aetion}
\CSLLeftMargin{6. }%
\CSLRightInline{Merola, D. \emph{et al.} The aetion coalition to advance
real-world evidence through randomized controlled trial emulation
initiative: oncology. \emph{Clinical Pharmacology \& Therapeutics}
\textbf{113}, 1217--1222 (2023).}

\bibitem[\citeproctext]{ref-merola2024calibrating}
\CSLLeftMargin{7. }%
\CSLRightInline{Merola, D. \emph{et al.} Calibrating observational
health record data against a randomized trial. \emph{JAMA Network Open}
\textbf{7}, e2436535--e2436535 (2024).}

\bibitem[\citeproctext]{ref-wang2023emulation}
\CSLLeftMargin{8. }%
\CSLRightInline{Wang, S. V. \emph{et al.}
\href{https://doi.org/10.1001/jama.2023.4221}{Emulation of randomized
clinical trials with nonrandomized database analyses: Results of 32
clinical trials}. \emph{Jama} \textbf{329}, 1376--1385 (2023).}

\bibitem[\citeproctext]{ref-heyard2024design}
\CSLLeftMargin{9. }%
\CSLRightInline{Heyard, R., Held, L., Schneeweiss, S. \& Wang, S. V.
\href{https://doi.org/10.1136/bmjmed-2023-000709}{Design differences and
variation in results between randomised trials and non-randomised
emulations: Meta-analysis of RCT-DUPLICATE data}. \emph{BMJ medicine}
\textbf{3}, (2024).}

\bibitem[\citeproctext]{ref-encoreFDA}
\CSLLeftMargin{10. }%
\CSLRightInline{Calibrating real-world evidence studies in oncology
against randomized trials: ENCORE (last accessed 11/28/2024). (2024).at
\textless{}\url{https://www.fda.gov/about-fda/oncology-center-excellence/calibrating-real-world-evidence-studies-oncology-against-randomized-trials-encore}\textgreater{}}

\bibitem[\citeproctext]{ref-rivera2024oncology}
\CSLLeftMargin{11. }%
\CSLRightInline{Rivera, D. R. \emph{et al.} The oncology QCARD
initiative: Fostering efficient evaluation of initial real-world data
proposals. \emph{Pharmacoepidemiology and Drug Safety} \textbf{33},
e5818 (2024).}

\bibitem[\citeproctext]{ref-gatto2022structured}
\CSLLeftMargin{12. }%
\CSLRightInline{Gatto, N. M. \emph{et al.} The structured process to
identify fit-for-purpose data: A data feasibility assessment framework.
\emph{Clinical Pharmacology \& Therapeutics} \textbf{111}, 122--134
(2022).}

\bibitem[\citeproctext]{ref-tasneem2012database}
\CSLLeftMargin{13. }%
\CSLRightInline{Tasneem, A. \emph{et al.} The database for aggregate
analysis of ClinicalTrials. Gov (AACT) and subsequent regrouping by
clinical specialty. \emph{PloS one} \textbf{7}, e33677 (2012).}

\bibitem[\citeproctext]{ref-eisenhauer2009new}
\CSLLeftMargin{14. }%
\CSLRightInline{Eisenhauer, E. A. \emph{et al.} New response evaluation
criteria in solid tumours: Revised RECIST guideline (version 1.1).
\emph{European journal of cancer} \textbf{45}, 228--247 (2009).}

\bibitem[\citeproctext]{ref-ThanCCR}
\CSLLeftMargin{15. }%
\CSLRightInline{Ton, T. G. N. \emph{et al.}
\href{https://doi.org/10.1158/1078-0432.CCR-22-0471}{Replication of
overall survival, progression-free survival, and overall response in
chemotherapy arms of nonâsmall cell lung cancer trials using real-world
data}. \emph{Clinical Cancer Research} \textbf{28}, 2844--2853 (2022).}

\bibitem[\citeproctext]{ref-mckelvey2024evaluation}
\CSLLeftMargin{16. }%
\CSLRightInline{McKelvey, B. A. \emph{et al.} Evaluation of real-world
tumor response derived from electronic health record data sources: A
feasibility analysis in patients with metastatic non--small cell lung
cancer treated with chemotherapy. \emph{JCO Clinical Cancer Informatics}
\textbf{8}, e2400091 (2024).}

\bibitem[\citeproctext]{ref-rwdRECIST}
\CSLLeftMargin{17. }%
\CSLRightInline{Chen, L. \emph{et al.}
\href{https://doi.org/10.1200/JCO.2023.41.16/_suppl.e21194}{Comparison
of response from RECIST1.1 and abstraction in real-world patients with
lung cancer.} \emph{Journal of Clinical Oncology} \textbf{41},
e21194--e21194 (2023).}

\bibitem[\citeproctext]{ref-rivera2022friends}
\CSLLeftMargin{18. }%
\CSLRightInline{Rivera, D. R. \emph{et al.} The friends of cancer
research real-world data collaboration pilot 2.0: Methodological
recommendations from oncology case studies. \emph{Clinical Pharmacology
\& Therapeutics} \textbf{111}, 283--292 (2022).}

\bibitem[\citeproctext]{ref-ackerman2024measurement}
\CSLLeftMargin{19. }%
\CSLRightInline{Ackerman, B. \emph{et al.} Measurement error and bias in
real-world oncology endpoints when constructing external control arms.
\emph{Frontiers in Drug Safety and Regulation} \textbf{4}, 1423493
(2024).}

\bibitem[\citeproctext]{ref-nikolakopoulou2014interpret}
\CSLLeftMargin{20. }%
\CSLRightInline{Nikolakopoulou, A., Mavridis, D. \& Salanti, G. How to
interpret meta-analysis models: Fixed effect and random effects
meta-analyses. \emph{BMJ Ment Health} \textbf{17}, 64--64 (2014).}

\bibitem[\citeproctext]{ref-wang2022harmonized}
\CSLLeftMargin{21. }%
\CSLRightInline{Wang, S. V. \emph{et al.}
\href{https://doi.org/10.1016/j.jval.2022.09.001}{HARmonized protocol
template to enhance reproducibility of hypothesis evaluating real-world
evidence studies on treatment effects: A good practices report of a
joint ISPE/ISPOR task force}. \emph{Value in Health} \textbf{25},
1663--1672 (2022).}

\bibitem[\citeproctext]{ref-guidelinegeneral}
\CSLLeftMargin{22. }%
\CSLRightInline{ICH guideline. General principles on plan, design and
analysis of pharmacoepidemiological studies that utilize real-world data
for safety assessment of medicines M14. Available from:
Https://database.ich.org/sites/default/files/ICH\_M14\_Step3\_DraftGuideline\_2024\_0521.pdf
(last accessed 01/07/2025). }

\bibitem[\citeproctext]{ref-hernan2022target}
\CSLLeftMargin{23. }%
\CSLRightInline{Hernán, M. A., Wang, W. \& Leaf, D. E. Target trial
emulation: A framework for causal inference from observational data.
\emph{Jama} \textbf{328}, 2446--2447 (2022).}

\bibitem[\citeproctext]{ref-hernan2016specifying}
\CSLLeftMargin{24. }%
\CSLRightInline{Hernán, M. A., Sauer, B. C., Hernández-Dı́az, S., Platt,
R. \& Shrier, I.
\href{https://doi.org/10.1016/j.jclinepi.2016.04.014}{Specifying a
target trial prevents immortal time bias and other self-inflicted
injuries in observational analyses}. \emph{Journal of clinical
epidemiology} \textbf{79}, 70--75 (2016).}
